### 2016-12-06 local-capability-machine.tex
g1 lemma

  \item[Store register contents to environment] Store the contents of each of the registers $r_1,\dots,r_n$ to the newly allocated memory
  \item[Allocate memory for record with environment capability, code capability, and acitvation code] Allocate a new piece of memory with room for a capability for the environment

@@

  \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the acitvation code and restrict it to an enter capability.

@@

\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
\label{lem:correctness-f1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    m & \defeq \hs_{f1} \uplus 
               \hs_\flag \uplus                
               \ms_{\var{link}} \uplus 
               \hs_\adv \uplus 
               \ms_{\malloc} \uplus 
               \hs_{\var{frame}} 
  \end{align*}
where 
\begin{align*}
  &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\dom(\ms_\link) = [\link,\link+1]\\
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
\end{align*}
and
\begin{itemize}
\item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
\item $\ms_\flag = [flag \mapsto 0]$
\item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
\item $\hs_\adv$ contains a global read-only capability for $\hs_\link$ on its first address. The remaining cells of the memory segment only contain instructions.
\end{itemize}
if 
  \[
    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
  \]
then
\[
  m'(\flag) = 0
\]  
\end{lemma}

@@

\begin{lemma}[Correctness lemma for \texttt{f2}]
  \label{lem:correctness-f2}
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\olf,\mathtt{2f},\mathtt{f2}) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{align*}
    &\dom(\hs_{f2}) = [\mathtt{f2}-\olf,\mathtt{2f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{f2}(\mathtt{f2}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f2}(\mathtt{f2}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f2}$ contains the code of $f2$.
  \item $\ms_\flag = [flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
\begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
  Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
  \begin{multline*}
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
    (\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
  \end{multline*}
Where 
\[
  m''  \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_\link \uplus 
        \hs_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk' \uplus
        \ms_{\var{frame}}
\]
Only the stack part of memory has changed. The stack and register-file at this point of execution is illustrated in figure~\ref{fig:f2-stack-before-adv}, where
\begin{itemize}
\item The range of authority for 
  \begin{itemize}
  \item $c_\stk'$ is $[\start_\stk,\addrend_\stk]$ (the entire stack)
  \item $c_\stk''$ is $[\start_\stk'',\addrend_\stk]$ (the unused part of the stack)
  \end{itemize}
  the two stack capabilities have the same permissions as $c_\stk$
\item $c_{f2}'$ points to the return address in f2's code.
\item $c_{\var{prp}}$ is a local enter-capability (with authority of all of the stack).
\item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
\end{itemize}

@@

\begin{lemma}[Correctness lemma for \texttt{f3}]
    \label{lem:correctness-f3}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f3} & \defeq ((\rwx,\glob),\mathtt{f3}-\olf,\mathtt{3f},\mathtt{f3}) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{f3} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{align*}
    &\dom(\hs_{f3}) = [\mathtt{f3}-\olf,\mathtt{3f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{f3}(\mathtt{f3}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f3}(\mathtt{f3}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f3}$ contains the code of $f3$.
  \item $\ms_\flag = [flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv$ contains a read-write capability for $\hs_\link$ on its first address and otherwise it consists only of code.
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}

@@

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
g = fun _ => let x = 0 in
               fun f =>
                 x := 0;
                 f();
                 x := 1;
                 f();
                 assert(x == 1)
\end{verbatim}

>>>>

  \item[Store register contents to environment] Store the contents of each of the registers $r_1,\dots,r_n$ to the newly allocated memory.
  \item[Allocate memory for record with environment capability, code capability, and acitvation code] Allocate a new piece of memory with room for a capability for the environment.

@@

  \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the acitvation code and restrict it to a global enter-capability.

@@

\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
\label{lem:correctness-f1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    m & \defeq \hs_{f1} \uplus 
               \hs_\flag \uplus                
               \ms_{\var{link}} \uplus 
               \hs_\adv \uplus 
               \ms_{\malloc} \uplus 
               \hs_{\var{frame}} 
  \end{align*}
where 
\begin{align*}
  &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\dom(\ms_\link) = [\link,\link+1]\\
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
\end{align*}
and
\begin{itemize}
\item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
\item $\ms_\flag = [\flag \mapsto 0]$
\item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
\item $\hs_\adv$ contains a global read-only capability for $\hs_\link$ on its first address. The remaining cells of the memory segment only contain instructions.
\end{itemize}
if 
  \[
    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
  \]
then
\[
  m'(\flag) = 0
\]  
\end{lemma}

@@

\begin{lemma}[Correctness lemma for \texttt{f2}]
  \label{lem:correctness-f2}
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\olf,\mathtt{2f},\mathtt{f2}) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{align*}
    &\dom(\hs_{f2}) = [\mathtt{f2}-\olf,\mathtt{2f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{f2}(\mathtt{f2}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f2}(\mathtt{f2}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f2}$ contains the code of $f2$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
\begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
  Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
  \begin{multline*}
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
    (\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
  \end{multline*}
Where 
\[
  m''  \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_\link \uplus 
        \hs_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk' \uplus
        \ms_{\var{frame}}
\]
Only the stack part of memory has changed. The stack and register-file at this point of execution is illustrated in figure~\ref{fig:f2-stack-before-adv}, where
\begin{itemize}
\item The range of authority for 
  \begin{itemize}
  \item $c_\stk'$ is $[\start_\stk,\addrend_\stk]$ (the entire stack)
  \item $c_\stk''$ is $[\start_\stk'',\addrend_\stk]$ (the unused part of the stack)
  \end{itemize}
  the two stack capabilities have the same permissions as $c_\stk$
\item $c_{f2}'$ points to the return address in f2's code.
\item $c_{\var{prp}}$ is a local enter-capability (with authority of all of the stack).
\item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
\end{itemize}

@@

\begin{lemma}[Correctness lemma for \texttt{f3}]
    \label{lem:correctness-f3}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f3} & \defeq ((\rwx,\glob),\mathtt{f3}-\olf,\mathtt{3f},\mathtt{f3}) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{f3} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{align*}
    &\dom(\hs_{f3}) = [\mathtt{f3}-\olf,\mathtt{3f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{f3}(\mathtt{f3}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f3}(\mathtt{f3}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f3}$ contains the code of $f3$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv$ contains a read-write capability for $\hs_\link$ on its first address and otherwise it consists only of code.
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}

@@

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
g = fun _ => let x = 0 in
               fun f =>
                 x := 0;
                 f();
                 x := 1;
                 f();
                 assert(x == 1)
\end{verbatim}
\begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$  $\dots$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
1g:  jmp $r_0$
f4:  reqglob $r1$
     reqperm $r_\stk$ $\encodePerm(\rwlx)$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0]$)
     load $r_1$ x
     assert $r_1$ 1
     cclear $\RegName \setminus \{r_0,\pcreg\}$
4f:  jmp $r_0$
\end{lstlisting}
Where the $\dots$ is the appropriate offset to make the capability point to \texttt{f4}.

\begin{lemma}[Correctness of $g1$]
      \label{lem:correctness-g1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{g1} & \defeq ((\entry,\local),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
    c_\stk & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{g1} \uplus 
        \ms_\flag \uplus                
        \ms_\link \uplus 
        \ms_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{align*}
    &\dom(\hs_{g1}) = [\mathtt{g1}-\olf,\mathtt{4f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{g1}(\mathtt{g1}-\olf) = ((\readonly,\glob),\link,\link,\link)$, $\ms_{g1}(\mathtt{g1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{g1}$ contains the code of $g1$ immediately followed by the code of $f4$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\link \mapsto c_\malloc]$
  \item $\hs_\adv$ contains a read-write capability for $\hs_\link$ on its first address and otherwise it consists only of instruction.
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
\lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$}
